---
title: "April Temp Analysis"
format: html
editor: visual
---

```{r}
#| load-libraries
#| echo: false
#| warning: false

library(tidyverse)
library(here)
library(dataRetrieval)
library(janitor)
```

```{r}
#| label: load-data
#| echo: false

temp.data <- read.csv(here("river_temp_data.csv"))
```

# Preparing Data for Analysis
```{r}
#| label: clean-data
#| echo: true

april.data <- temp.data|>
  select(agency_cd,site_no,temp_C,Date,month,year)|>
  filter(month=="Apr")|>
  clean_names()
```

## Monthly Temperature Trends in April
```{r}
#| label: glm
#| echo: true

april.lm <- glm(temp_c ~ year*as.factor(site_no), data=april.data)
summary(april.lm)
```
From the results of this General Linear Model, there is a slight positive trend in April river temperature across years at sites 14166000 and 14211720 and a slight negative trend in April river temperature change across years at site 14211010. However, these changes are insignificant as all relevant p-values are less than 0.05. We therefore fail to reject the null hypothesis that there is no change in April river temperature across years and sites in this area.
