---
title: "october data analysis"
format: html
editor: visual
---

load packages
```{r}
library(tidyverse)
library(here)
```

load data
```{r}
temp_data <- read.csv(here("river_temp_data.csv"))
```


fit GLM model
```{r}
# filter for July temperatures
oct_data <- subset(temp_data, month == "Oct")
oct_data$site_no <- as.factor(oct_data$site_no)
# fit a GLM model to see if there is a trend in July temperatures over time, separated by site
oct_model <- glm(temp_C ~ year*site_no, data = oct_data)
summary(oct_model)
```
At sites 14166000 and 14211010, there is no significant trend October water 
temperatures across years. Both sites have a p-value over a significance
level of 0.1 (0.558 and 0.151, respectively.) At site 14211720, 
there could be a significant positive relationship between year and 
October water temperatures. This is evidenced by the p-value of 0.066, 
which is less than 0.1. Based on the GLM, we would expect a 0.054 degree
Celcius increase in water temperature per year. 

Summary Statistics
```{r}
summary <- oct_data|>
  select(agency_cd, site_no, temp_C, Date, month, year)|>
  group_by(site_no)|>
  summarise(mean_temp_c=mean(temp_C), max_temp_c=max(temp_C), min_temp_c=min(temp_C))

print(summary)
```

Site 14211720 has the highest water temperatures on average and 
Site 1421101 had the lowest water temperatures on average. 
