---
title: "summary_qmd"
format: html
editor: visual
---

```{r}
#| label: libraries
#| echo: false
#| warning: false

library(tidyverse)
library(here)
library(ggplot2)
library(patchwork)
```


```{r}
#| label: load-data
#| echo: true

temp_data <- read.csv(here("river_temp_data.csv"))
```

```{r}
#| label: summary
#| echo: true

summary <- temp_data|>
  select(agency_cd, site_no, temp_C, Date, month, year)|>
  group_by(site_no)|>
  summarise(mean_temp_c=mean(temp_C), max_temp_c=max(temp_C), min_temp_c=min(temp_C))

print(summary)
```

```{r}
#| label: october
#| echo: true

# filter for Otober temperatures
oct_data <- subset(temp_data, month == "Oct")
oct_data$site_no <- as.factor(oct_data$site_no)
# fit a GLM model to see if there is a trend in October temperatures over time, separated by site
oct_model <- glm(temp_C ~ year*site_no, data = oct_data)
summary(oct_model)
```

At sites 14166000 and 14211010, there is no significant trend in October water temperatures across years. Both sites have a p-value over a significance level of 0.1 (0.558 and 0.151, respectively.) At site 14211720, there could be a significant positive relationship between year and October water temperatures. This is evidenced by the p-value of 0.066, which is less than 0.1. Based on the GLM, we would expect a 0.054 degree Celcius increase in water temperature per year.

```{r}
#| label: april
#| echo: true

april.data <- temp.data|>
  select(agency_cd,site_no,temp_C,Date,month,year)|>
  filter(month=="Apr")|>
  clean_names()


april.lm <- glm(temp_c ~ year*as.factor(site_no), data=april.data)
summary(april.lm)
```

From the results of this General Linear Model, there is a slight positive trend in April river temperature across years at sites 14166000 and 14211720 and a slight negative trend in April river temperature change across years at site 14211010. However, these changes are insignificant as all relevant p-values are less than 0.05. We therefore fail to reject the null hypothesis that there is no change in April river temperature across years and sites in this area.

```{r}
#| label: plot
#| echo: false

oct_plot <- ggplot(oct_data, aes(x=site_no, y=temp_C))+
  geom_boxplot() +
  ylim(0,20)+
  labs(title="October",
       x="Site Number", y="Temperature (°C)") +
  theme_bw()

april_plot <- ggplot(april.data, aes(x=as.factor(site_no), y=temp_c))+
  geom_boxplot() +
  ylim(0,20)+
  labs(title="April", x="Site Number", y="Temperature (°C)")+
  theme_bw()

april_plot
oct_plot+april_plot
```
Across both April and October, only one site displayed a significant trend of water temperature across years. In both months, the GLM coefficients of sites 14166000 and 14211010 were all above the significance threshold of 0.1. Site 1421170 had a p-value under 0.1 during October only. There was evidence of a positive trend (0.05 degrees per year) in October water temperature across years at that site. Site 11421170 was the warmest on average across both April and October. Water temperatures were consistently warmer in October than April.